---
title: "Section 15 - Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 15 - Data Visualization

## Layers  
The principles of data visualization are built on the concept of layers.
Traditionally, every data visualization is made up of three core layers  
- **Data** (the underlying numerical data)  
- **Aesthetics** (the chart decoration such as gridlines, labels, etc)  
- **Geometries** (the physical manifestation of the data such as points on graph, lines, etc)  
```{r}
library(ggplot2)
```
```{r}
pl <- ggplot(data=mtcars, # <-- Data Layer
             aes(x=mpg, y=hp)) # <-- Aesthetics Layer
```
```{r}
pl + geom_point() # <-- Geometries Layer
```

The next three layers add additional information to make visualizations more understandable.  
- **Facets**  
- **Statistics**  
- **Coordinates**  
```{r}
pl + geom_point() + facet_grid(cyl ~.) # <-- Facets Layer
```

```{r}
pl + geom_point() + facet_grid(cyl ~.) +
  stat_smooth() # <-- Statistics Layer
```

```{r}
pl + geom_point() + facet_grid(cyl ~.) + stat_smooth() + 
  coord_cartesian(xlim=c(15,25)) # <-- Coordinates Layer
```


The 7th and final layer is added for tweaking visual appearance  
- **Theme** (font size, colors, background, etc)  

```{r}
pl + geom_point() + facet_grid(cyl ~.) + stat_smooth() + coord_cartesian(xlim=c(15,25)) +
  theme_bw() # <-- Theme Layer
```

```{r}
library(ggplot2movies)
```

### [R Studio Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)

### Histograms

```{r}
colnames(movies)
data <- ggplot(movies, aes(x=rating)) # Data Layer
geometry <- data + 
  geom_histogram(binwidth = 0.1, color='red', fill='pink', alpha=0.4) # Geometry Layer
aesthetics <- geometry + xlab('Movie Rating') + ylab('Count') # Aesthetics Layer
titled <- aesthetics + ggtitle('Movie Ratings')
titled
altfill <- data + geom_histogram(binwidth = 0.1, aes(fill=..count..))
altfill
```

### Scatterplot

```{r}
df <- mtcars
pl <- ggplot(df, aes(x=wt, y=mpg))
pl + geom_point(size=5, alpha=0.7)
pl + geom_point(aes(size=hp), color='red')
pl + geom_point(size=5, alpha=0.7, aes(shape=factor(cyl), color=factor(cyl)))
```


```{r}
pl + geom_point(aes(color=hp), size=5) + scale_color_gradient(low='green', high='red')
```


### Barplot

```{r}
df <- mpg
head(mpg)
pl <- ggplot(df, aes(x=class))
pl + geom_bar(aes(fill=drv))
```

```{r}
pl + geom_bar(aes(fill=drv), position='dodge')
pl + geom_bar(aes(fill=drv), position='fill')
```


### Boxplots

```{r}
df <- mtcars
head(df)

pl <- ggplot(df, aes(x=cyl, y=mpg))
pl + geom_boxplot() # ERROR
pl <- ggplot(df, aes(x=factor(cyl), y=mpg))
pl + geom_boxplot()
```

Box plots are already very information dense so best not to add too many layers to them  

```{r}
pl + geom_boxplot(aes(fill=factor(cyl))) + theme_bw()
```


### 2-Variable Plots
```{r}
pl <- ggplot(movies, aes(x=year, y=rating))
pl + geom_bin2d() + scale_fill_gradient(high='red', low='navy')
```


```{r}
pl + geom_hex()
```

```{r}
```
```{r}
pl + geom_hex()
```

```{r}
pl + geom_density2d()
```


### Coordinates and Faceting
```{r}
pl <- ggplot(mpg, aes(x=displ, y=hwy)) + geom_point()
pl
pl + coord_cartesian(xlim=c(1, 4), ylim=c(15,30))
pl + coord_fixed(ratio = 1/3)
```

```{r}
help("facet_grid")

pl + facet_grid(. ~ cyl)
pl + facet_grid(cyl ~ drv)
```


### Themes
```{r}
pl <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point()
pl
```

```{r}
theme_set(theme_minimal()) # sets default theme for all plots
pl
```

```{r}
pl + theme_dark()
```

More themes!!!
```{r}
library(ggthemes)
```


```{r}
pl + theme_economist()
```




## Exercises

### Histogram of hwy mpg values:  
```{r}
ggplot(mpg, aes(x=hwy)) + geom_histogram(fill='red', alpha=0.5)
```

### Barplot of car counts per manufacturer with color fill defined by cyl count

```{r}
ggplot(mpg, aes(x=manufacturer)) + geom_bar(aes(fill=factor(cyl)))
```

### Switch now to use the txhousing dataset that comes with ggplot2

###Create a scatterplot of volume versus sales. Afterwards play around with alpha and color arguments to clarify information.

```{r}
ggplot(txhousing, aes(x=sales, y=volume)) + 
  geom_point(color='blue', fill='blue', alpha=0.5)
```


### Add a smooth fit line to the scatterplot from above. Hint: You may need to look up geom_smooth()

```{r}
ggplot(txhousing, aes(x=sales, y=volume)) + 
  geom_point(color='blue', fill='blue', alpha=0.5) +
  geom_smooth(color='red')
```


## Plotly Visualizations!

### [Getting Started with Plotly in R](https://plot.ly/R/getting-started)

Find more docs [here](https://plot.ly/ggplot2/)

```{r}
library(plotly)
```


```{r}
pl <- ggplot(mtcars, aes(mpg, wt)) + geom_point()
pl
```


```{r}
gpl <- ggplotly(pl)
gpl
```












