---
title: "Section 13 - Advanced Programming in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 13 - Advanced Programming in R

## Built-In Functions
`seq()` will generate a sequence of continuous numbers starting at `start`, ending at `end` inclusive and moving `by` places at a time.  
e.g. `seq(0, 10, by = 2)` will print `[  0  2  4  6  8  10  ]`

`sort()` will sort the values in a vector.  
```
v <- c(1, 4, 7, 2, 13, 3, 11)
sort(v)

[  1  2  3  4  7  11  13  ]
```

you can also sort descending with `sort(v, decreasing = TRUE)`  
sorting also works with alpha values case insensitive


`rev(v)` will reverse the order of a vector

`str(df)` and `summary(df)` give stats on dataframes

`append(v1, v2)` will attach the values of v2 onto the end of v1 (not sure how this is different from `c(v1, v2)`)

Data type can be checked with the `is.datatype()` function. e.g.  
```
v <- c(1,2,3)
is.vector(v) => TRUE
is.data.frame(v) => FALSE
```

Data types can be converted (if compatible) using the `as.datatype()` function. e.g.  
```
v <- c(1,2,3)
as.list(v) => List representation of v:
[[1]]
[1]  1

[[2]]
[1]  2

[[3]]
[1]  3
```

## Apply
The apply function will apply a passed function to elements of an itterable (like a vector or list)

`sample(x = range, n)` will return `n` random samples within the given `range` e.g.  
`sample(x = 1:100, 3)` might return `[  77  13  92  ]`

```
v <- c(1,2,3,4,5)
addrand <- function(x) {
  ran <- sample(1:100, 1)
  return (x + ran)
}
lapply(v, addrand)
```
lapply will return a list (could look like:)  
```
[[1]]
[1]  12

[[2]]
[1]  79

[[3]]
[1]  64

[[4]]
[1]  8

[[5]]
[1]  92
```

for a more simple result, `sapply()` can be used to return a vector result e.g.  
```
sapply(v)
[1]  33 101 71 27 85
```

One more example:
```
v <- 1:5

times2 <- function(num) {
  return(num * 2)
}
print(v)
result <- sapply(v, times2)
print(result)
```
Will print:  
```
[1]  1 2 3 4 5
[1]  2 4 6 8 10
```

`help(sapply)` can show other variants of apply function


## Anonymous Functions
```
v <- 1:5
print(sapply(v, function(num){num * 2}))
```
Will acheive the same result as  
```
v <- 1:5
dub <- function(num) {
 return(num * 2)
}

print(sapply(v, dub))
```

### Using apply with multiple inputs
suppose a function has been defined with 2 or more inputs like  
```
v <- 1:5
add.num <- function(x, num) {
  return(x + num)
}
print(sapply(v, add.num))
```
Will yield an error. Apply will pass the itterator's element to the first argument, but any additional non-optional arguments must be specified like so:  
```
print(sapply(v, add.num, num = 5))
```

## Math Functions
`abs(x)` returns the absolute value of x  
`sum(v)` returns sum of vector  
`round(x, n)` rounds x to n digits

More functions [referenced here](https://cran.r-project.org/doc/contrib/Short-refcard.pdf)


## Regex
`grepl(r)` returns a logical  
`grep(r)` returns an index  

```{r}
text <- "Hi there, do you know who you are voting for?"
text
grepl('voting', text)
grepl('dog', text)
```

```{r}
v <- c("a", "b", "c", "d", "d")
v
grepl("b", v)

grep("d", v)
grep("e", v)
```


## Date and Timestamps
```{r}
today <- Sys.Date()
today
class(today)

c <- "1990-01-01"
c
class(c)
my.date <- as.Date(c)
my.date
class(my.date)

# as.Date("Nov-03-1990")
parsed.date <- as.Date("Nov-03-1990", format = "%b-%d-%y")
parsed.date
```

Date codes for defining format:  
`help(strptime)`


Times use the POSIXct type  
```{r}
as.POSIXct("11:02:03", format = "%H:%M:%S")
```

But more common, the `strptime()` function will be used  
```{r}
strptime("11:02:03", format="%H:%M:%S")
```

