---
title: "Section 8 - Matrices"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 8 - Matrices

### Creating a matrix
```{r}
v <- 1:10
v
matrix(v)

matrix(v, nrow=2)
matrix(1:12, byrow=F, nrow=4)
matrix(1:12, byrow=F, nrow=4)
```

```{r}
goog <- c(450, 451, 452, 445, 468)
msft <- c(230, 231, 232, 233, 220)

stocks <- c(goog, msft)

stock.matrix <- matrix(stocks, byrow=T, nrow=2)

days <- c("Mon", "Tue", "Wed", "Thu", "Fri")
st.names <- c("GOOG", "MSFT")
colnames(stock.matrix) <- days
rownames(stock.matrix) <- st.names
```


### Matrix Arithmetic
```{r}
mat <- matrix(1:25, byrow=T, nrow=5)
mat + 2
mat / 2
mat ^ 2

mat > 15

mat[mat > 15]
```

```{r}
colSums(stock.matrix)
rowSums(stock.matrix)

rowMeans(stock.matrix)
colMeans(stock.matrix)
```


```{r}
FB <- c(111, 112, 113, 120, 145)

tech.stocks <- rbind(stock.matrix, FB)
tech.stocks

avg <- rowMeans(tech.stocks)
avg
tech.stocks <- cbind(tech.stocks, avg)
tech.stocks
```


### Matrix selection and indexing
```{r}
mat <- matrix(1:50, byrow=T, nrow=5)
mat

mat[1, ]
mat[, 1]
mat[1:3, ]

mat[1:2, 1:3]
```


### Matrix factors
```{r}
animal <- c("d","c","d","c","c")
id <- c(1,2,3,4,5)

#Nominal factors (no inherent order)
fact.aml <- factor(animal)

#Ordinal factors (can be assigned <, > comparisons)
temps <- c("cold", "med", "hot", "hot","hot", "cold", "med")
fact.temp <- factor(temps, ordered=T, levels=c("cold", "med", "hot"))
fact.temp
summary(fact.temp)
summary(temps)
```



# Exercises
### Create 2 vectors A and B, where A is (1,2,3) and B is (4,5,6). With these vectors, use the cbind() or rbind() function to create a 2 by 3 matrix from the vectors. You'll need to figure out which of these binding functions is the correct choice.
```{r}
A <- c(1,2,3)
B <- c(4,5,6)
rbind(A, B)
```

### Ex 2: Create a 3 by 3 matrix consisting of the numbers 1-9. Create this matrix using the shortcut 1:9 and by specifying the nrow argument in the matrix() function call. Assign this matrix to the variable mat
```{r}
mat <- matrix(1:9, nrow=T)
```

### Ex 3: Confirm that mat is a matrix using is.matrix()
```{r}
is.matrix(mat)
```

### Ex 4: Create a 5 by 5 matrix consisting of the numbers 1-25 and assign it to the variable mat2. The top row should be the numbers 1-5.
```{r}
mat2 <- matrix(1:25, byrow=T, nrow=5)
mat2
```

### Ex 5: Using indexing notation, grab a sub-section of mat2 from the previous exercise that looks like this:
```{r}
mat2[2:3, 2:3]
```

### Ex 6: Using indexing notation, grab a sub-section of mat2 from the previous exercise that looks like this:
```{r}
mat2[4:5, 4:5]
```

### Ex 7: What is the sum of all the elements in mat2?
```{r}
sum(mat2)
```

### 
Ex 8: Ok time for our last exercise! Find out how to use runif() to create a 4 by 5 matrix consisting of 20 random numbers (4*5=20).
```{r}
matrix(runif(20)*100, byrow=T, nrow=4)
```









