---
title: "Section 9 - Data Frames"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 9 - Data Frames
## Basics
### Built-in dataframes
```{r}
state.x77
USPersonalExpenditure
women
WorldPhones
data()
```

### Constructing your own
```{r}
days <- c("Mon",  "Tue", "Wed", "Thu", "Fri")
temp <- c(22.2, 21, 23, 24.3, 25)
rain <- c(T, T, F, F, T)
df <- data.frame(days, temp, rain)
df
str(df)
summary(df)
```

## Indexing and Selecting
```{r}
df[1, ]
df[ ,1]

df[ , "rain"]
df[1:5 , c("days", "temp")]
df$temp
```

Selection with '$' returns all the values for that columns  
Selection with ["colname"] returns a single column dataframe


## Subsetting
```{r}
subset(df, subset = rain==T)
subset(df, subset= temp>23)
```


```{r}
sorted.temp <- order(df["temp"])
sorted.temp

df[sorted.temp,]
```
```{r}
desc.temp <- order(-df$temp)
desc.temp

df[desc.temp,]
```


## Overview of DF Operations
```{r}
c1 <- 1:10
c2 <- letters[1:10]
c1
c2

df <- data.frame(col.name.1=c1, col.name.2=c2)
df
```

### I/O
`write.csv(dataframe, file="path/to/file.csv")` (this also saves index as a new column)
`d2 <- read.csv("path/to/file.csv")`
```{r}
write.csv(df, file="saved_df.csv")
df2 <- read.csv("saved_df.csv")
df2
```


```{r}
df
nrow(df)
ncol(df)
colnames(df)
rownames(df)
str(df)
summary(df)

df[[5, 2]]
df[[5, "col.name.2"]]

df[[2, "col.name.1"]] <- 9999
df
```

```{r}
mtcars
head(mtcars)

#All the same selection
mtcars$mpg        #vector of values
mtcars[ , "mpg"]  #vector of values
mtcars[ , 1]      #vector of values
mtcars[["mpg"]]   #vector of values

mtcars["mpg"]     #dataframe
mtcars[1]         #dataframe
```

### Adding Rows
```{r}
df2 <- data.frame(col.name.1 = 2000, col.name.2 = "new")
df2
df

dfnew <- rbind(df, df2)
dfnew
```

### Adding Columns
```{r}
df$newCol <- 2 * df$col.name.1
df
```

```{r}
df$newcol.copy <- df$newCol
head(df)
```


### Rename columns
```{r}
colnames(df)
colnames(df) <- c("1", "2", "3", "4")
head(df)

colnames(df)[1] <- "NEW COL NAME"
head(df)
```


### Select multiple rows
```{r}
df[1:10, ]
df[1:3, ]
head(df, 3)
df[-2, ] #All but row 2
```

```{r}
head(mtcars)

mtcars[mtcars$mpg > 20 , ] #important to include comma (to specify cols returned)
```

```{r}
head(mtcars)

mtcars[ (mtcars$mpg > 20) & (mtcars$cyl == 6) , ]

mtcars[ (mtcars$mpg > 20) & (mtcars$cyl == 6) , c("mpg", "cyl", "hp")] #or...
subset(mtcars, mpg > 20 & cyl == 6)
```

### Dealing with missing data
```{r}
is.na(mtcars) # => dataframe with is.na called on each value

any(is.na(df)) # => single boolean value, true if any entry in df is NA
```

Replace all NA values with 0:
```{r}
df[is.na(df)] <- 0
```
Or replace NA's in a particular column with that column's average
```{r}
mtcars$mpg[is.na(mtcars$mpg)] <- mean(mtcars$mpg)
```




# Exercises

### Ex 1: Recreate the following dataframe by creating vectors and using the data.frame function:

```{r}
Name <- c("Sam", "Frank", "Amy")
Age <- c(22,25,26)
Weight <- c(150, 165, 120)
Sex <- c("M", "M", "F")

data.frame(row.names = Name, Age, Weight, Sex)
```

### Ex 2: Check if mtcars is a dataframe using is.data.frame()
```{r}
is.data.frame(mtcars)
```

### Ex 3: Use as.data.frame() to convert a matrix into a dataframe
```{r}
mat <- matrix(1:25, nrow=5)

data.frame(mat)
```


### Ex 4: Set the built-in data frame mtcars as a variable mf. We'll use this df variable for the rest of the exercises.
```{r}
mf <- mtcars
```

### Ex 5: Display the first 6 rows of mf
```{r}
head(mf)
```

### Ex 6: What is the average mpg value for all the cars?
```{r}
mean(mf$mpg)
```

### Ex 7: Select the rows where all cars have 6 cylinders (cyl column)
```{r}
subset(mf, cyl == 6)
```


### Ex 8: Select the columns am,gear, and carb.
```{r}
mf[c("am", "gear", "carb")]
```

### Ex 9: Create a new column called performance, which is calculated by hp/wt.
```{r}
mf$performance <- (mf$hp / mf$wt)
mf$performance
```

### Ex 10: Your performance column will have several decimal place precision. Figure out how to use round() (check help(round)) to reduce this accuracy to only 2 decimal places.
```{r}
mf$performance <- round(mf$performance, 2)
mf$performance
```

### Ex 10: What is the average mpg for cars that have more than 100 hp AND a wt value of more than 2.5.
```{r}
mean(subset(mf, hp > 100 & wt > 2.5)$mpg)
```

### Ex 11: What is the mpg of the Hornet Sportabout?
```{r}
mf["Hornet Sportabout", "mpg"]
```



