---
title: "Section 7 - R Basics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 8 - R Basics

## Data Types
class(variable) can be used to check the type of variable
### Numeric
+ float
+ integer

### Logical
+ FALSE or F
+ TRUE or T

### Characters
use single or double quotes  
"double quotes" are more common

### Vectors
created with combine function  
  e.g. `c(1,2,3,4,5)` or `c(1:5)`  
all elements in a vector must be of the same class (auto-conversion)  
  e.g.`c("USA", 20, 30)` => `["USA", "20", "30"]` and `c(FALSE, 20, 30)` => `[0, 20, 30]`

Vectors can be named with the name function  
  e.g. `names(vector) <- c(name1, name2, ...)`

#### Vector Functions
+ standard arithmetic performed element-wise c(1,2,3) + c(4,5,6) = [5,7,9]
+ sum(vector) returns a sum over all elements `sum(c(1,2,3))` => 6
+ mean(vector)
+ sd(vector)
+ min(vector)
+ max(vector)
+ prod(vector)

#### Vector selections
```{r}
v1 <- c(100,200,300)
v2 <- c("a", "b", "c")

# Index starts at 1 in R...
v1[2]
v2[3]

v1[c(1,2)]
v2[c(1,3)]
```

```{r}
v <- c(1,2,3,4,5,6,7,8,9,10)
# start and stop index are both inclusive
v[2:4]
v[7:10]

v <- c(1,2,3,4)
names(v) <- c('a', 'b', 'c', 'd')
v
v[2]
v['b']
v[c('c','d','a')]

v[v>2]

v>2

my.filter <- v>2
v[my.filter]
```


## Getting help in R
help('topic')

??topic

help.search('topic)




# R Basics Exercise

### What is two to the power of 5?
```{r}
2 ** 5
```

### Create a vector called stock.prices with the following data points: 23,27,23,21,34
```{r}
stock.prices <- c(23,27,23,21,34)
stock.prices
```

### Assign names to the price data points relating to the day of the week, starting with Mon, Tue, Wed, etc...
```{r}
names(stock.prices) <- c("Mon","Tue","Wed","Thu","Fri")
stock.prices
```

### Create a vector called over.23 consisting of logicals that correspond to the days where stock price was more than $23
```{r}
over.23 <- stock.prices > 23
```

### Use the over.23 vector to filter out the stock.prices vector and only return the day and prices where the price was over $23
```{r}
stock.prices[over.23]
```


### Use a built in function to find the day the price was the highest
```{r}
max(stock.prices) 
max.price <- stock.prices == max(stock.prices)
stock.prices[max.price]
```

`1:10` is a shorthand way to create a vector of numbers 1 to 10





