---
title: "Section 14 - Data Manipulation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 14 - Data Manipulation
```{r}
# install.packages('dplyr')
# install.packages('nycflights13')

library(dplyr)
library(nycflights13)
```

## DPLYR

- filter will perform a simple subset on a dataframe, accepting as many filters as you need to pass as additional arguments
- slice will return a 'slice' of rows from a dataframe
- arrange will reorder rows in a dataframe based on a specified column (mult columns can be specified)
- select will just return the specified columns of dataframe
- rename allows easy renaming of columns in dataframe (`rename(df, new.name = old.name)`)
- select will return all the unique values in a column
- mutate will add a new column that can be built using function of existing functions (returns entire dataframe)
- transmute will return only the new column created
- summarise will aggregate the data in a dataframe using some function like mean
- sample_n will return a random n number of rows from dataframe
- sample_frac will return a portion of rows from the dataframe
```{r}
head(flights)
head(filter(flights, month==11, day==3, carrier=="AA"))
slice(flights, 1:10)
head(arrange(flights, year, month, day, desc(arr_time)))
head(select(flights, carrier, arr_delay))
head(rename(flights, airline_carrier = carrier))
distinct(select(flights, carrier))
head(mutate(flights, new_col = arr_delay - dep_delay))
summarise(flights, avg_air_time=mean(air_time, na.rm = TRUE))
sample_n(flights, 10)
```


## The Pipe Operator
```{r}
df <- mtcars
result <- arrange(sample_n(filter(df, mpg>20), size=5), desc(mpg))
print(result)
```

Above code is hard to read due to extensive nesting

```{r}
a <- filter(df, mpg>20)
b <- sample_n(a, size=5)
result <- arrange(b, desc(mpg))
print(result)
```

This code is not great either with so many unnecessary variable assignments

```{r}
result <- df %>% filter(mpg>20) %>% sample_n(size=5) %>% arrange(desc(mpg)) 
print(result)
```

^ Nice clean andefficient code using pipe operator

## TIDYR
```{r}
# install.packages('tidyr')
# install.packages('data.table')
```

```{r}
library(tidyr)
library(data.table)

comp <- c(1,1,1,2,2,2,3,3,3)
yr <- c(1998,1999,2000,1998,1999,2000,1998,1999,2000)
q1 <- c(runif(9, min=0, max=100))
q2 <- c(runif(9, min=0, max=100))
q3 <- c(runif(9, min=0, max=100))
q4 <- c(runif(9, min=0, max=100))

df <- data.frame(comp=comp, year=yr, Qtr1=q1, Qtr2=q2, Qtr3=q3, Qtr4=q4)
df
```


### gather(data, key, value, columns_to_gater)  
This function will take "wide" dataframes and condense several columns into one.  
In this example, Qtr1, Qtr2, etc. are not actually variables, but instead just different
values that the "Quarter" variable might take on. Therefore, these columns should be
condensed into a single "Quarter" column
 
> Well-formatted tables will have a single row for each observation and a single row for each variable. Every column should reveal something new about that particular observation.
 
```{r}
gather(df, Quarter, Revenue, Qtr1:Qtr4)
```
 
```{r}
stocks <- data.frame(
  time = as.Date('2009-01-01') + 0:9,
  x = rnorm(10,0,1),
  y = rnorm(10,0,2),
  z = rnorm(10,0,4)
)
head(stocks)
```

```{r}
stocks.gathered <- stocks %>% gather(stock, price, x, y, z)
head(stocks.gathered)
```

But what if we want to spread that data back out?

### spread()

```{r}
stocks.gathered %>% spread(stock, price)
```

But notice that this is really not well-formatted or "tidy" data


### separate(data, col, into, sep)  
This function will split out combined data based on `sep` in a column (`col`) into several columns identified in `into` argument
```{r}
df <- data.frame(new.col=c(NA, "a.x", "b.y", "c.z"))
df
```

```{r}
separate(df, new.col, c('ABC', 'XYZ'))
```

```{r}
df <- data.frame(new.col=c(NA, "a-x", "b-y", "c-z"))
df
```
```{r}
separate(data=df, col=new.col, into=c("abc", "xyz"), sep='-')
```

But of course there is a reverse of this as well...

### unite(data, new.col, [list,of,cols], sep)  
This will unite data from several columns joined by `sep` in `list,of,cols` into a single column named `new.col`

```{r}
df.sep <- separate(data=df, col=new.col, into=c("abc", "xyz"), sep='-')
df.sep

unite(df.sep, new.joined.col, abc, xyz)
```




## DPLYR Exercises:

### Return rows of cars that have an mpg value greater than 20 and 6 cylinders.
```{r}
filter(mtcars, mpg>20, cyl==6)
```

### Reorder the Data Frame by cyl first, then by descending wt.

```{r}
arrange(mtcars, cyl, desc(wt))
```

### Select the columns mpg and hp
```{r}
select(mtcars, mpg, hp)
```

### Select the distinct values of the gear column.
```{r}
distinct(select(mtcars, gear))
```

### Create a new column called "Performance" which is calculated by hp divided by wt.
```{r}
mutate(mtcars, Performance = hp / wt)
```

### Find the mean mpg value using dplyr.
```{r}
summarise(mtcars, avg_mg = mean(mpg))
```

### Use pipe operators to get the mean hp value for cars with 6 cylinders.
```{r}
mtcars %>% filter(cyl == 6) %>% summarise(avg_hp = mean(hp))
```




